---
title: "EDA"
author: "shoili"
date: "April 8, 2017"
output: pdf_document
---

```{r, message=FALSE}
setwd("~/Documents/airbnb_MLproj")
library(dplyr)
library(lubridate)
```

This document takes a look at the dataset. Makes the necessary joins and outlines steps to be taken for feature generation before the prediction model is built. 

```{r}
# read data
calendar <- read.csv("calendar.csv", strip.white = T, stringsAsFactors = F)
head(calendar)
listings <- read.csv("listings.csv", strip.white = T, stringsAsFactors = F)
# head(listings)
reviews <- read.csv("reviews.csv", strip.white = T, stringsAsFactors = F)
# head(reviews)
```

Calendar contains the prices for all available days so we strip out the unavailable days data. We then convert the price to a number from a character field. 

```{r}
# clean calendar table
table(calendar$available)

available_listings <- calendar %>%
  filter(available == "t")

available_listings$price_num <- as.numeric(sub("\\$","", available_listings$price))
summary(available_listings$price_num)

available_listings <- available_listings[!is.na(available_listings$price_num), ]

rm(calendar)
```
Just read in available_listings.csv from now on instead of calendar. 
This file has 2898 unique listings. The dates of the listings range from 2016-09-06 to 2017-09-05. We need to generate what month/week of year the listing is for to identify seasonal trends. (done)

```{r}
available_listings <- available_listings %>%
  mutate(month = month(date), week = isoweek(date))

write.csv(available_listings, file = "available_listings.csv")
```

The new field 'price_num' is the response variable for our model. Now we need to join the data from listings and reviews. Taking a look at the data contained in these. 

```{r}
colnames(listings)
# head(listings)
```

After feature generation, join listings with available_listings by columns 'id' and 'listing_id'. 

Possible features - 

1. listing_url remove
2. scrape_id, last_scrape remove
3. name, summary, space, description. ??
4. experiences offered seems all None so remove. 
5. neighborhood overview, notes, access, interation ??
6. transit. 1/0 for field empty or no? not empty probably means easy access to public transit. (check this assumption)
7. house rules. no smoking check as feature?
8. thumbnail, picture etc url. 1/0 feature. if no picture that's probably bad
9. host_id. retain this i guess some hosts suck
10. host location and name prolly not important
11. host_about ??
12. host_response_time. "N/A"                "within an hour"     "within a few hours" "within a day"       "a few days or more" convert values into categorical numeric variable
13. host_acceptance_rate, host_response_rate. turn into numeric. 
14. host_is_superhost. convert t/f to 1/0
15. more urls. 1/0 for exist or no?
16. host_neighborhood. meh
17. host_listings_count and next column are same. 
18. host_verifications. put count?
19. host profile pic and identity verified convert to numeric from t/f
20. street ??
21. neighborhood cleaned. dont use neighborhood. numeric encode?
22. city state zipcode country lat long
23. property_type and room_type. convert to numeric categorical variable
24. accomodates bathrooms bedrooms beds keep as is. (numeric)
25. bed_type. categorical numeric
26. amenities. 1/0 for internet, cable etc
27. square feet. keep as is numeric
28. remove all the prices and keep as possible alternate responses probably
29. min/max night
30. requires license, instant bookable, require guest things. convert all to 1/0 from t/f
31. cancellation policy. convert to numeric categorical
32 review_Scores_* check what these mean

```{r}
colnames(reviews)
```
Reviews can also be joined to available_listings by listing_id. 

Possible features - 

1. Number of reviews a listing has. (done)
2. Avg length of review. (chars done. also have words?)
3. Avg sentiment of review.
4. fraction of reviews positive
5. fraction of reviews negative

```{r}
review_features <- reviews %>%
  mutate(reviewlen = nchar(comments)) %>%
  group_by(listing_id) %>%
  mutate(reviewcount = n(), avgreviewlen = round(mean(reviewlen)))
```

Need to train sentiment analysis. Use python maybe?